High-Dimensional Regularized Discriminant Analysis

### 2013-05-10 reports/paper-hdrda.Rnw
Removed most of the references to Friedman's RDA paper. Also, added the classifier's computational complexity.

A reliable supervised classification of tumors from gene expression microarray
data is essential for successful diagnosis and treatment of diseases such as
cancer; however, the analysis of microarray data often presents difficult
challenges because the number of gene expression levels is typically much
greater than the sample size. Consequently, several sparse and regularized
classifiers have been proposed, often with strict covariance assumptions or
eigenvalue adjustments to ensure that the classifiers are well-posed. Here, we
propose a novel modification of the regularized discriminant analysis classifier
popularized by \cite{Friedman:1989tm}. Our proposed classifier provides a
straightforward interpretation and employs a new dimension-reduction technique
to enable a computationally efficient model-selection procedure, thus overcoming
the computational shortcomings of \citeauthor{Friedman:1989tm}'s classifier for
high-dimensional data. Moreover, we demonstrate with four high-dimensional
microarray data sets that our proposed classifier is competitive with and often
superior to several recently proposed sparse and regularized classifiers in
terms of classification accuracy. Finally, we provide an implementation of our
proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.

@@

Alternatively, several researchers, including \cite*{Zhang:2010va},
\cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te},
\cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed
regularization methods to improve the estimation of the population covariance
matrices before applying either linear or quadratic discriminant
analysis. Covariance-matrix regularization methods often are based on the
well-known ridge-regression approach of \cite{Hoerl:1970cd} and adjust the
eigenvalues of the sample covariance matrices to ensure positive
definiteness. For excellent overviews of the regularization problem in
discriminant analysis, see \cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.

Here, we propose an improvement to the popular regularized discriminant
analysis classifier of \cite{Friedman:1989tm}, who has proposed pooling of the
covariance-matrix estimators from linear and quadratic discriminant analysis and
has additionally incorporated a covariance-matrix shrinkage term to ensure that
the classifier is well-posed. \citeauthor{Friedman:1989tm}'s classifier can
attain improved classification accuracy, compared to linear and quadratic
discriminant analysis when $p < N$. However, in its original form,
\citeauthor{Friedman:1989tm}'s classifier is impractical when $p \gg N$ because
the decision rule requires the calculation of the inverses and determinants of
large $p \times p$ covariance-matrix estimators for each class. Furthermore,
\citeauthor{Friedman:1989tm}'s classifier has two tuning parameters that are
selected via cross-validation, which adds a substantial computational burden to
the model-selection process.

In this paper we modify \citeauthor{Friedman:1989tm}'s classifier by first
constructing the pooling matrix as a convex combination of the covariance-matrix
estimators employed in linear and quadratic discriminant analysis and
demonstrate that this simplification improves the interpretation of the pooling
operation of \citeauthor{Friedman:1989tm}'s classifier. Given that the
assumption of equal population covariance matrices for linear discriminant
analysis is often too stringent in practice \citep{Clemmensen:2011kr}, our
proposed classifier can relax the linearity assumption by instead incorporating
unequal covariance matrices while retaining linear and quadratic discriminant
analysis as special cases. We then establish the high-dimensional regularized
discriminant analysis (\emph{HDRDA}) classifier by applying
covariance-matrix regularization with a parameterization that includes a large
family of shrinkage methods, such as those from \cite{Srivastava:2007ww},
\cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
\cite{Xu:2009fl}.

We show that our proposed classifier is invariant to any adjustments of the
eigenvalues corresponding to the null space of the pooled sample covariance
matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
decision rule of the \emph{HDRDA} classifier and show that the sum related to
the null space of the pooled sample covariance matrix can be discarded. Hence,
no gain in classification performance can be obtained by adjusting these zero
eigenvalues. This invariance property addresses the computational shortcomings
of \citeauthor{Friedman:1989tm}'s classifier and provides a substantial increase
in computational efficiency and model selection. In fact, we show that the
inverse and determinants can be reduced to calculations of $n_k \times n_k$
matrices, where $n_k$ is the sample size of the $k$th class $(k = 1, \ldots,
K)$. This reduction in calculation is clearly advantageous because $n_k$ is
often quite small in microarray data sets (e.g., 15--20 observations).

@@

Quadratic discriminant analysis is the optimal Bayesian decision rule with
respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the probability
density function of the multivariate normal distribution with known mean vector
$\bm \mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm
\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where
$\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$
positive-definite matrices. Because $\bm \mu_k$ and $\bm \Sigma_k$ are typically
unknown, we assign an unlabeled observation $\bm x$ to class $\omega_k$ with the
sample quadratic discriminant analysis classifier
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k} (\bm x - \xbar_k)^{T}\widehat{\bm
    \Sigma}_k^{-1}(\bm x - \xbar_k) + \log |\widehat{\bm \Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum likelihood
estimators for $\bm \mu_k$ and $\bm \Sigma_k$, respectively, and $\bm A^{T}$
denotes the matrix transpose of $\bm A \in \mathbb{R}_{m \times p}$. If we
assume further that $\bm \Sigma_k = \bm \Sigma$ $(k = 1, \ldots, K)$, then the
pooled sample covariance matrix $\widehat{\bm \Sigma}$ is substituted for
$\widehat{ \bm \Sigma}_k$ in \eqref{eq:qda}, where
\begin{align}
	\widehat{\bm \Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\bm \Sigma}_k \label{eq:pooled-cov}
\end{align}
is the maximum likelihood estimator for $\bm \Sigma$. In this case \eqref{eq:qda}
reduces to the sample linear discriminant analysis classifier, and the
log-determinant is omitted because it is constant across all classes.

@@

\subsection{Friedman's Regularized Discriminant Analysis Classifier}

\cite{Friedman:1989tm} has proposed regularized discriminant analysis by
incorporating a weighted average of the sample covariance matrix $\widehat{ \bm
  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm \Sigma}$ to
estimate $\bm \Sigma_k$ with
\begin{align}
  \widehat{\bm \Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad (k =
  1, \ldots, K),\label{eq:sig-lambda}
\end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
\bm S_k = n_k \widehat{\bm \Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $ \bm
S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. We can interpret
\eqref{eq:sig-lambda} as a covariance-matrix estimator that borrows from
$\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to better estimate $ \bm
\Sigma_k$. If $\lambda = 0$, \eqref{eq:sig-lambda} is equal to $\widehat{\bm
  \Sigma}_k$ used in \eqref{eq:qda}. Alternatively, if $\lambda = 1$,
\eqref{eq:sig-lambda} reduces to \eqref{eq:pooled-cov}.

To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
covariance-matrix estimator
\begin{align}
	\widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
\end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Thus, the \emph{pooling} parameter
$\lambda$ controls the amount borrowed from $\widehat{\bm \Sigma}$ to estimate
$\bm \Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the
amount of eigenvalue shrinkage applied. Finally, by substituting
\eqref{eq:sig-rda} into \eqref{eq:qda}, we have the regularized discriminant
analysis classifier
\begin{align}
	D_{RDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\widehat{\bm \Sigma}_k(\lambda, \gamma)^{-1}(\bm x - \xbar_k)  + \log |\widehat{\bm \Sigma}_k(\lambda, \gamma)|. \label{eq:rda}
\end{align}
If $\gamma$ is sufficiently small, \eqref{eq:sig-rda} can be singular or
near-singular. For this case \cite{Friedman:1989tm} has suggested replacing the
zero eigenvalues of \eqref{eq:sig-rda} with a small positive constant that
permits a numerically stable matrix inversion.

@@

Here, we define our adaptation of \citeauthor{Friedman:1989tm}'s classifier by
first simplifying the formulation of $\widehat{\bm \Sigma}_k(\lambda)$ in
\eqref{eq:sig-lambda} and demonstrating its clear interpretation as a linear
combination of the crossproducts of the training observations centered by their
respective class sample means. We define the convex combination

@@

where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. Thus, we avoid the
convex combination of $n_k$ and $N$ as defined by \cite{Friedman:1989tm}. By
rewriting \eqref{eq:sig-lambda-alternative} in terms of the observations $
\bm x_i$ $(i = 1, \ldots, N)$, each centered by its class sample mean, we attain a
clear interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,

@@

where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
each of the $N$ observations in estimating $\bm \Sigma_k$. Hence, for $0 <
\lambda \le 1$, we estimate $\bm \Sigma_k$ with $N$ rather than $n_k$
observations. Below, we will show that the pooling operation is advantageous in
increasing the rank of each $\widehat{\bm \Sigma}_k(\lambda)$ from
rank$(\widehat{\bm \Sigma}_k)$ to rank$(\widehat{\bm \Sigma})$ for $0 < \lambda
\le 1$. Notice that if $\lambda = 0$, then the observations from the remaining
$K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$,
corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the
weights in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to

@@

[\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm x_N^{T}]^{T}$ such
that $\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Similarly, for $\lambda = 0$, we
define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(0)} \bm
  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k = \bm X_k^{T} \bm X_k$. In
Figure \ref{fig:hdrda-contours} we plot the contours of five multivariate normal
populations for $\lambda = 0$ with unequal covariance matrices. As $\lambda$
approaches 1, the contours become more similar, resulting in identical contours
for $\lambda = 1$.

@@

proposed to increase eigenvalues (approximately) equal to 0. Here, we define the
eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as

@@

constant. The choice of $\alpha_k$ allows for a flexible formulation of

@@

if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to
\eqref{eq:sig-rda}. Now, substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda},
we define the high-dimensional regularized discriminant analysis classifier

@@

\subsection{Enhanced Computational Efficiency}

Next, we demonstrate that our proposed classifier in \eqref{eq:hdrda} lends
itself to a more efficient calculation. We decompose \eqref{eq:hdrda} into a sum
of two components, where the first summand consists of matrix operations applied
to low-dimensional matrices and the second summand corresponds to the null space
of $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant
quadratic forms across all classes and can be omitted. For $p \gg N$, the
constant component involves determinants and inverses of high-dimensional
matrices, and by ignoring these calculations, a substantial reduction in
computational costs is achieved relative to Friedman's classifier in
\eqref{eq:rda}. Furthermore, a byproduct is that adjustments to the associated
eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we utilize the singular
value decomposition to calculate efficiently the eigenvalue decomposition of
$\widehat{\bm \Sigma}$, further reducing the computational costs of the
\emph{HDRDA} classifier.

@@

Conveniently, we see that the inverse and determinant of $\bm W_k$ are each in
terms of the inverse and determinant of $\bm Q_k \in \mathbb{R}_{n_k \times
  n_k}$, respectively, as well as the diagonal matrix $\bm \Gamma_k \in
\mathbb{R}_{q \times q}$. This formulation is especially advantageous when the
number of classes $K$ is large because $n_k$ is then potentially much smaller
than $q$. The expressions given in Proposition \ref{proposition:hdrda-W_k} also
expedite the selection of $\lambda$ and $\gamma$ via cross-validation. The most
time-consuming matrix operations involved in computing $\bm W_k^{-1}$ and $|\bm
W_k|$ is $\bm X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent
of $\lambda$ and $\gamma$. Hence, we need calculate $\bm X_k \bm U_1$ only once
for each cross-validation fold. The subsequent operations in calculating $\bm
W_k^{-1}$ and $|\bm W_K|$ can be simply updated for different pairs of $\lambda$
and $\gamma$.

@@

We evaluated the classification performance of each classifier by randomly
partitioning the data such that $2/3$ of the observations were allocated as
training data and the remaining $1/3$ of the observations were allocated as a
test data set. For each random partition, we applied the feature-selection
method from \cite{Dudoit:2002ev} to the training data to choose the $d$ features
having the largest ratio of their between-group to within-group sums of squares.
We then filtered the test data set with the corresponding $d$ features. After
training each classifier from the filtered training data, we classified the
filtered test data sets and computed the proportion of mislabeled test
observations to estimate the classification error rate for each
classifier. Repeating this process 1000 times, we computed the average of the
error-rate estimates for each classifier. We remark that our simulation design
avoids the selection bias discussed in \cite{Ambroise:2002fa}. We did not
include \citeauthor{Friedman:1989tm}'s classifier in our studies to avoid its
excessive computations.

@@

In our simulation study, we compared our proposed \emph{HDRDA} classifier to
several recently proposed classifiers, including the Dudoit, Pang, Tong, and Guo
classifiers, each of which assumes independence of the features conditional on
class membership. Despite the rapid computational performance of diagonal
classifiers and the reduction in the number of parameters to estimate,
\cite{Mai:2012bf} and \cite{Fan:2012iq} have noted that diagonal classifiers can
often yield inferior classification performance to other classification
methods. We have shown that the \emph{HDRDA} classifier often yields superior
classification accuracy to the diagonal classifiers considered, confirming the
assertions of \cite{Mai:2012bf} and \cite{Fan:2012iq}. Furthermore, the
implementation of our classifier in the {\tt sparsediscrim} R package was
comparable to that of the diagonal classifiers in terms of computational
runtime. Moreover, when preliminary variable selection was applied to the data
in our classification study, the time to perform model selection with our
proposed classifier was nearly instantaneous.

>>>>

A reliable supervised classification of tumors from gene expression microarray
data is essential for successful diagnosis and treatment of diseases such as
cancer; however, because the number of gene expression levels measured is
typically much greater than the sample size, the analysis of microarray data
often presents difficult statistical and computational challenges. Here, we
propose a new regularized discriminant analysis classifier that incorporates
dimension reduction and covariance-matrix shrinkage to enable a computationally
efficient classifier and model-selection procedure. We demonstrate with four
high-dimensional microarray data sets that our proposed classifier is often
superior to several recently proposed sparse and regularized classifiers in
terms of classification accuracy. Finally, we provide an implementation of our
proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.

@@

Alternatively, several researchers, including \cite*{Zhang:2010va},
\cite{Ye:2009gd}, \cite{Ji:2008wp}, \cite*{Guo:2007te},
\cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed
regularization methods to improve the estimation of the population covariance
matrices before applying either linear or quadratic discriminant
analysis. Covariance-matrix regularization methods often are based on the
well-known ridge-regression approach of \cite{Hoerl:1970cd} and adjust the
eigenvalues of the sample covariance matrices to ensure positive
definiteness. For excellent overviews of the regularization problem in
discriminant analysis, see \cite{Murphy:2012uq}, \cite{Mkhadri:1997gy}, and
\cite{Friedman:1989tm}.

Here, we propose the high-dimensional regularized discriminant analysis
(\emph{HDRDA}) by first constructing a pooling matrix as a convex combination of
the covariance-matrix estimators employed in linear and quadratic discriminant
analysis. Given that the assumption of equal population covariance matrices for
linear discriminant analysis is often too stringent in practice
\citep{Clemmensen:2011kr}, our proposed classifier can relax the linearity
assumption by instead incorporating unequal covariance matrices while retaining
linear and quadratic discriminant analysis as special cases. We then establish
\emph{HDRDA} classifier by applying covariance-matrix regularization with a
parameterization that includes a large family of shrinkage methods, such as
those from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods
studied by \cite{Ramey:2013ji} and \cite{Xu:2009fl}. This regularization has the
effect of stabilizing the eigenvalues of the covariance-matrix estimators and
ensuring that the classifier is well-posed.

We show that our proposed classifier is invariant to any adjustments of the
eigenvalues corresponding to the null space of the pooled sample covariance
matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
decision rule of the \emph{HDRDA} classifier and show that the sum related to
the null space of the pooled sample covariance matrix can be discarded. Hence,
no gain in classification performance can be obtained by adjusting these zero
eigenvalues. This invariance property provides a substantial increase in
computational efficiency of model training and selection. In fact, we show that
the inverse and determinants can be reduced to calculations of $n_k \times n_k$
matrices, where $n_k$ is the sample size of the $k$th class $(k = 1, \ldots,
K)$. This reduction in calculation is clearly advantageous because $n_k$ is
often quite small in microarray data sets (e.g., 15--20 observations).

@@

Quadratic discriminant analysis is the optimal Bayesian decision rule with
respect to a $0-1$ loss function when $p(\bm x | \omega_k)$ is the probability
density function of the multivariate normal distribution with known mean vector
$\bm \mu_k \in \mathbb{R}_{p \times 1}$ and known covariance matrix $\bm
\Sigma_k \in \mathbb{R}_{p \times p}^{>}$, $k = 1, 2, \ldots, K$, where
$\mathbb{R}_{p \times p}^{>}$ denotes the cone of real $p \times p$
positive-definite matrices. Because $\bm \mu_k$ and $\bm \Sigma_k$ are typically
unknown, we assign an unlabeled observation $\bm x$ to class $\omega_k$ with the
sample quadratic discriminant analysis classifier
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k} (\bm x - \xbar_k)^{T}\widehat{\bm
    \Sigma}_k^{-1}(\bm x - \xbar_k) + \log |\widehat{\bm \Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum likelihood
estimators of $\bm \mu_k$ and $\bm \Sigma_k$, respectively, and $\bm A^{T}$
denotes the matrix transpose of $\bm A \in \mathbb{R}_{m \times p}$. If we
assume further that $\bm \Sigma_k = \bm \Sigma$ $(k = 1, \ldots, K)$, then the
pooled sample covariance matrix $\widehat{\bm \Sigma}$ is substituted for
$\widehat{ \bm \Sigma}_k$ in \eqref{eq:qda}, where
\begin{align}
	\widehat{\bm \Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\bm \Sigma}_k \label{eq:pooled-cov}
\end{align}
is the maximum likelihood estimator for $\bm \Sigma$. In this case \eqref{eq:qda}
reduces to the sample linear discriminant analysis classifier, and the
log-determinant is omitted because it is constant across all classes.

@@

@@

Here, we define the \emph{HDRDA} classifier by first formulating the
covariance-matrix estimator $\widehat{\bm \Sigma}_k(\lambda)$ and demonstrating
its clear interpretation as a linear combination of the crossproducts of the
training observations centered by their respective class sample means. We define
the convex combination

@@

where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $ \bm x_i$ $(i =
1, \ldots, N)$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,

@@

where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$. We
can interpret \eqref{eq:sig-lambda-alternative2} as a covariance-matrix
estimator that borrows from $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to
better estimate $ \bm \Sigma_k$. From \eqref{eq:sig-lambda-alternative2},
$\lambda$ weights the contribution of each of the $N$ observations in estimating
$\bm \Sigma_k$. Hence, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with
$N$ rather than $n_k$ observations. Below, we will show that the pooling
operation is advantageous in increasing the rank of each $\widehat{\bm
  \Sigma}_k(\lambda)$ from rank$(\widehat{\bm \Sigma}_k)$ to rank$(\widehat{\bm
  \Sigma})$ for $0 < \lambda \le 1$. Notice that if $\lambda = 0$, then the
observations from the remaining $K - 1$ classes do not contribute to the
estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm
  \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights in
\eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to

@@

      [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm x_N^{T}]^{T}$
      such that $\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Similarly, for
      $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots,
        \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k =
      \bm X_k^{T} \bm X_k$. In Figure \ref{fig:hdrda-contours} we plot the
      contours of five multivariate normal populations for $\lambda = 0$ with
      unequal covariance matrices. As $\lambda$ approaches 1, the contours
      become more similar, resulting in identical contours for $\lambda = 1$.

@@

proposed to increase eigenvalues (approximately) equal to 0. To further improve
the estimation of $\bm \Sigma_k$ and to stabilize the estimator's inverse, we
define the eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as

@@

constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount
borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm \Sigma_k$, and the
\emph{shrinkage} parameter $\gamma$ determines the amount of eigenvalue
shrinkage applied. The choice of $\alpha_k$ allows for a flexible formulation of

@@

if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to the
\emph{RDA} classifier from \cite{Friedman:1989tm}. Now, substituting
\eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the high-dimensional
regularized discriminant analysis classifier

@@

\subsection{Enhanced Computational Efficiency}
\label{sec:computational-efficiency}

Next, we demonstrate that our proposed classifier in \eqref{eq:hdrda} lends
itself to a more efficient calculation. We decompose \eqref{eq:hdrda} into a sum
of two components, where the first summand consists of matrix operations applied
to low-dimensional matrices and the second summand corresponds to the null space
of $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix
operations performed on the null space of $\widehat{\bm \Sigma}$ yield constant
quadratic forms across all classes and can be omitted. For $p \gg N$, the
constant component involves determinants and inverses of high-dimensional
matrices, and by ignoring these calculations, a substantial reduction in
computational costs is achieved. Furthermore, a byproduct is that adjustments to
the associated eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we
utilize the singular value decomposition to calculate efficiently the eigenvalue
decomposition of $\widehat{\bm \Sigma}$, further reducing the computational
costs of the \emph{HDRDA} classifier.

@@

Conveniently, we see that the inverse and determinant of $\bm W_k$ are each in
terms of the inverse and determinant of $\bm Q_k \in \mathbb{R}_{n_k \times
  n_k}$, respectively, as well as the diagonal matrix $\bm \Gamma_k \in
\mathbb{R}_{q \times q}$. Hence, the resulting formulation requires $O(n_k p q)$
calculations to compute both $\bm W_k^{-1}$ and $| \bm W_k |$ and is especially
advantageous when the number of classes $K$ is large because $n_k$ is then
potentially much smaller than $q$. The expressions given in Proposition
\ref{proposition:hdrda-W_k} also expedite the selection of $\lambda$ and
$\gamma$ via cross-validation because the most time-consuming matrix operation
involved in computing $\bm W_k^{-1}$ and $|\bm W_k|$ is $\bm X_k \bm U_1 \in
\mathbb{R}_{n_k \times q}$, which is independent of $\lambda$ and
$\gamma$. Therefore, we need calculate $\bm X_k \bm U_1$ only once for each
cross-validation fold. The subsequent operations in calculating $\bm W_k^{-1}$
and $|\bm W_K|$ can be simply updated for different pairs of $\lambda$ and
$\gamma$.

@@

We evaluated the classification performance of each classifier by randomly
partitioning the data such that $2/3$ of the observations were allocated as
training data and the remaining $1/3$ of the observations were allocated as a
test data set. For each random partition, we applied the feature-selection
method from \cite{Dudoit:2002ev} to the training data to choose the $d$ features
having the largest ratio of their between-group to within-group sums of squares.
We then filtered the test data set with the corresponding $d$ features. After
training each classifier from the filtered training data, we classified the
filtered test data sets and computed the proportion of mislabeled test
observations to estimate the classification error rate for each
classifier. Repeating this process 1000 times, we computed the average of the
error-rate estimates for each classifier. We remark that our simulation design
avoids the selection bias discussed in \cite{Ambroise:2002fa}.

@@

In our simulation study, we compared our proposed \emph{HDRDA} classifier to
several recently proposed classifiers, including the Dudoit, Pang, Tong, and Guo
classifiers, each of which assumes independence of the features conditional on
class membership. Despite the rapid computational performance of diagonal
classifiers and the reduction in the number of parameters to estimate,
\cite{Mai:2012bf} and \cite{Fan:2012iq} have noted that diagonal classifiers can
often yield inferior classification performance to other classification
methods. We have shown that the \emph{HDRDA} classifier often yields superior
classification accuracy to the diagonal classifiers considered, confirming the
assertions of \cite{Mai:2012bf} and \cite{Fan:2012iq}. Furthermore, the
implementation of our classifier in the {\tt sparsediscrim} R package was
comparable to that of the diagonal classifiers in terms of computational
runtime. Moreover, when preliminary variable selection was applied to the data
in our classification study, the time to perform model selection with our
proposed classifier was nearly instantaneous. Hence, our proposed \emph{HDRDA}
classifier is competitive in terms of classification accuracy for modern
small-sample, high-dimensional data sets and requires substantially less
computing time than standard regularized classifiers, such as regularized
discriminant analysis proposed by \cite{Friedman:1989tm}.
